% "chapters/intro.tex"

%	$Id: intro.tex,v 1.9 2005/12/13 04:14:42 fang Exp $

\chapter{Introduction}
\label{sec:intro}

% a better quote:
\begin{quotation}
But in our enthusiasm, we could not resist a radical overhaul of the
system, in which all of its major weaknesses have been exposed,
analyzed, and replaced with new weaknesses.
\end{quotation}

\begin{flushright}
{\itshape -- Bruce Leverett, ``Register Allocation in Optimizing Compilers''}
\end{flushright}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the beginning, there was \index{CSP}CSP\cite{ref:csp}.

This document describes the language specification for the \hac\ Language.  

Should be largely implementation-independent.  
But we discuss some of the issues.  

Our implementation is a multi-phase compiler with four phases:
\begin{itemize}
\item Compile -- like precompiling modules
\item Link
\item Unroll -- heirarchical expansion of top-level instantiations.  
\item Finalize -- \emph{unique} instantiation after connections.  
\end{itemize}

Compile-time is ...

Link-time is ...

Unroll-time is ...

Finalize is ...


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Purpose}
\label{sec:intro:purpose}

What is the purpose of \hac?
What is the meaning of life?

What this IS:
\begin{itemize}
\item a hardware description language
\end{itemize}

What this IS NOT:
\begin{itemize}
\item a sequential programming language
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Roots}
\label{sec:intro:roots}
% or History

The \hac\ language is based on the CAST\index{CAST}
(Caltech Asynchronous Synthesis Tool) language.  

Discuss limitations.  
Hierarchical information was lost, as a result of 
flattening identifiers into strings.  

% The fact that this is named \hac\ implies that there is or was a 
% predecessor named ART.  :P
% (The fact that this is named \hac\ is a purely jab at the original ART, 
% implying that this version is in some ways superior.)

Religious differences, much like the
\verb|vi| versus \verb|emacs| holy war, 
led to a divergence of implementations and interpretations.  


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
\label{sec:intro:overview}

In Chapter blah we cover blah.  
In Chapter foo we cover foo.  

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

